Training a Gaming Agent on Brainwaves

### 2020-10-20 ErrP.tex
Correcciones minimas

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of Event-Related Potential (ERP) elicited by a person attending a recognizable error.  These Electroencephalographic (EEG)  signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their brain signals are captured. The game consists of a grid, where a green spot has to reach a desired target in the fewest amount of steps.  Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the mechanics of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate low classification accuracies and a limited scope of transfer learning schemes.
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

The experimental procedure is summarized in Figure \ref{diag:complete_flow}. The proposed system has two distinct parts. This first part consists of the collection of brainwave signals from a person that is watching an agent play a game.  The agent knows the game rules but not how to win it. The second part, the gaming agent learning phase, is where the agent can learn the winning strategy using the person's feedback to improve its own performance.

@@

Once the headset is correctly applied, the OpenVibe Acquisition Server program, from the OpenVibe platform~\cite{OPEN-VIBE-PAPER}, is launched and configured with a sampling rate of $250$~Hz. A $50$~Hz notch filter is applied to filter out power line noise. An additional bandpass filter between $0.5$~Hz and $60$~Hz is applied. Data are handled and processed with the OpenVibe Designer, from the same platform, using 8 channels for the brain data (one channel per electrode) and an additional channel to record the stimulus, which corresponds to a game movement performed by the agent.  After everything is connected, the subject is seated in a comfortable chair in front of a computer screen. The brightness of the screen is set to the maximum setting to avoid any visual inconvenience in which the subject can not distinguish the components of the game that appear on the screen.

@@

In this way, the raw data of a brainwave session are processed into an array of matches where each element is an array of epochs tagged with a number specifying the prediction of the classifier, i.e. if the epoch corresponds to an action that made the agent move further from the goal (hit) or an action that made the agent moves closer to the goal (no-hit). The ErrP is expected to be found in hits. To get the data ready for classification, the stimulus channel is removed to classify the signals using only the EEG data. Each epoch is regularized using a MinMaxScaler, i.e. subtracting the minimum value in the epoch and dividing by the signal peak-to-peak amplitude~\cite{Zhou2019}.  The eight channels are concatenated using the MNE Vectorizer function, which transforms the data matrix into a single array sample. Lastly, these data are used by the classification module as information to train and test a classifier. Five different classification algorithms are used:  Logistic Regression, Multilayer Perceptron with a hidden layer of 100 neurons (i.e. default values for the Scikit-Learn MLPClassifier), Random Forest, KNeighbours with k=3 and finally a linear kernel Support Vector Classifier (i.e. SVM)~\cite{Lotte2018}.

@@

Complementary, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier.  Each point corresponds to a run session of 200 game play repetitions.  The y-value represent the average number of steps the agent takes to reach the goal using a specific Q-Table for a run session. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

In the case for OHC 5 and 6, the reward information obtained from the brainwaves is not enough to train the agent effectively. Figures \ref{fig:avg_steps} for OHC 5 and 6 show no apparent learning, as the amount of steps to reach the goal doesn't decrease when trained.  These results are also consistent with the classification ROC curves, shown in Figures~\ref{fig:rocsubjects}, where the area under the curve for OHCs 5 and 6 are close to chance level.  Both OHCs have less recorded data from the sessions in comparison to the rest of the OHCs.  This variation in performance for different OHCs has been studied extensively in BCI~\cite{Chavarriaga2014}.  Besides low data samples, there are other reasons affecting the classification accuracy:  cognitive reasons (i.e. the OHC not paying extensively attention to the game dynamics), very low SNR of the ErrP component or even the BCI-illiteracy phenomena where the specific OHC's signals do not contain the expected component response~\cite{Yousefi2019}.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  A classifier can be trained to recognize Error Potential from observational human critics that observe the agent playing the game.  The gaming agent plays randomly and movements are marked triggered by the identification of an error potential from the human critic.  Those movements are used as rewards in a Reinforcement Learning scheme to train a Q-Table.  The gaming agent plays the game again, but this time the number of required steps to solve the game is reduced.  If rewards are provided based on random signals,  no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data show that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic.

@@

At the same time, effective agent training depends on the OHC's training data. Results confirm the futility or complexity of using Transfer Learning~\cite{Wu2016}: training a classifier with data obtained from one OHC, and using the same classifier to identify ErrPs for another OHC does not increases the performance of the agent.  Despite that, the rewards generated from different subject's classifiers can be used to train the same Q-Table to improve its performance, which may lead to strategies where the overall performance is enhanced based on the information from different human critics at the same time.  We hypothesize that as long as there are accurate rewards obtained from high specificity classifiers, there is extra information that can be used by the agent to get an improved Q-Table.  Additionally, it seems to be an agreement in terms of the subjective interpretation of what may be an appropriate movement to reach the goal, and different OHCs produce rewards for the same type of movements.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of Event-Related Potential (ERP) elicited by a person attending a recognizable error.  These Electroencephalographic (EEG)  signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their brain signals are captured. The game consists of a grid, where a blue spot has to reach a desired target in the fewest amount of steps.  Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the mechanics of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate low classification accuracies and a limited scope of transfer learning schemes.
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

The experimental procedure is summarized in Figure \ref{diag:complete_flow}. The proposed system has two distinct parts. This first part consists of the collection of brainwave signals from an OHC that is watching an agent play a game.  The agent knows the game rules but not how to win it. The second part, the gaming agent learning phase, is where the agent can learn the winning strategy using the OHC's feedback to improve its own performance.

@@

Once the headset is correctly applied, the OpenVibe Acquisition Server program, from the OpenVibe platform~\cite{OPEN-VIBE-PAPER}, is launched and configured with a sampling rate of $250$~Hz. A $50$~Hz notch filter is applied to filter out power line noise. An additional bandpass filter between $0.5$~Hz and $60$~Hz is applied. Data are handled and processed with the OpenVibe Designer, from the same platform, using 8 channels for the brain data (one channel per electrode) and an additional channel to record the stimulus, which corresponds to a game movement performed by the agent.  After everything is connected, the subject is seated in a comfortable chair in front of a computer screen. The brightness of the screen is set to the maximum setting to avoid any visual inconvenience in which the subject cannot distinguish the components of the game that appear on the screen.

@@

In this way, the raw data of a brainwave session are processed into an array of matches where each element is an array of epochs tagged with a number specifying the prediction of the classifier, i.e. if the epoch corresponds to an action that made the agent move further from the goal (hit) or an action that made the agent move closer to the goal (no-hit). The ErrP is expected to be found in hits. To get the data ready for classification, the stimulus channel is removed to classify the signals using only the EEG data. Each epoch is regularized using a MinMaxScaler, i.e. subtracting the minimum value in the epoch and dividing by the signal peak-to-peak amplitude~\cite{Zhou2019}.  The eight channels are concatenated using the MNE Vectorizer function, which transforms the data matrix into a single array sample. Lastly, these data are used by the classification module as information to train and test a classifier. Five different classification algorithms are used:  Logistic Regression, Multilayer Perceptron with a hidden layer of 100 neurons (i.e. default values for the Scikit-Learn MLPClassifier), Random Forest, KNeighbours with k=3 and finally a linear kernel Support Vector Classifier (i.e. SVM)~\cite{Lotte2018}.

@@

Complementary, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier.  Each point corresponds to a run session of 200 game play repetitions.  The y-values represent the average number of steps the agent takes to reach the goal using a specific Q-Table for a run session. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

In the case for OHC 5 and 6, the reward information obtained from the brainwaves is not enough to train the agent effectively. Figures \ref{fig:avg_steps} for OHC 5 and 6 show no apparent learning, as the amount of steps to reach the goal doesn't decrease when trained.  These results are also consistent with the classification ROC curves, shown in Figures~\ref{fig:rocsubjects}, where the area under the curve for OHCs 5 and 6 are close to chance level.  Both OHCs have less recorded data from the sessions in comparison to the rest of the OHCs.  This variation in performance for different OHCs has been studied extensively in BCI~\cite{Chavarriaga2014}.  Besides low data samples, there are other reasons affecting the classification accuracy:  cognitive reasons (i.e. the OHC not paying extensive attention to the game dynamics), very low SNR of the ErrP component or even the BCI-illiteracy phenomena where the specific OHC's signals do not contain the expected component response~\cite{Yousefi2019}.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  A classifier can be trained to recognize Error Potential from OHCs that observe the agent playing the game.  The gaming agent plays randomly and movements are marked triggered by the identification of an error potential from the OHC.  Those movements are used as rewards in a Reinforcement Learning scheme to train a Q-Table.  The gaming agent plays the game again, but this time the number of required steps to solve the game is reduced.  If rewards are provided based on random signals,  no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data show that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the OHC.

@@

At the same time, effective agent training depends on the OHC's training data. Results confirm the futility or complexity of using Transfer Learning~\cite{Wu2016}: training a classifier with data obtained from one OHC, and using the same classifier to identify ErrPs for another OHC does not increase the performance of the agent.  Despite that, the rewards generated from different subject's classifiers can be used to train the same Q-Table to improve its performance, which may lead to strategies where the overall performance is enhanced based on the information from different OHCs at the same time.  We hypothesize that as long as there are accurate rewards obtained from high specificity classifiers, there is extra information that can be used by the agent to get an improved Q-Table.  Additionally, it seems to be an agreement in terms of the subjective interpretation of what may be an appropriate movement to reach the goal, and different OHCs produce rewards for the same type of movements.
